---
title: "Ridge, Lasso, Elastic Net, SCAD and MCP with gwrs"
author: "Bahadir Yuzbasi"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Ridge, Lasso, Elastic Net, SCAD and MCP with gwrs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(gwrs)
```

## Scope and interfaces

This guide covers geographically weighted ridge, lasso, elastic-net, SCAD and
MCP regression. All five estimators fit a local intercept and local slopes.
Ridge shrinks coefficients; Lasso, Elastic Net, SCAD and MCP can also set local
slopes to zero. SCAD and MCP use nonconvex penalties, so convergence does not
by itself establish a global optimum or recovery of the true active set.

```{r interface-table, echo=FALSE}
knitr::kable(data.frame(
  Estimator = c("GWR-Ridge", "GWR-Lasso", "GWR-Elastic Net", "GWR-SCAD", "GWR-MCP"),
  Formula = c("gwr_ridge()", "gwr_lasso()", "gwr_en()", "gwr_scad()", "gwr_mcp()"),
  Matrix = c("gwr_ridge_fit()", "gwr_lasso_fit()", "gwr_en_fit()", "gwr_scad_fit()", "gwr_mcp_fit()")
))
```

## Run a small example

The installed example constructs a deterministic grid on the unit square,
uses the first 48 locations for training and the last 16 for prediction,
and fits all five estimators. Coordinates and distances are dimensionless.
The example makes no statistical comparison or performance claim.

The demonstration fixes the adaptive bisquare neighborhood at 18 training
neighbors. It uses penalty 0.08 for the initial fits, Elastic Net mixing 0.5,
and the default SCAD/MCP concavity constants 3.7 and 3. Spatial CV then uses
three common folds (seed 42) and the explicit penalty grid 0.2, 0.08, 0.02.
These are example settings, not the settings used in the manuscript.

```{r run-example}
example_file <- system.file("examples", "five-methods.R", package = "gwrs")
source(example_file, local = knitr::knit_global())
```

Inspect the selected penalty for each method:

```{r selected-penalties}
lapply(cv, function(z) z$best)
```

Inspect SCAD coefficients and new-location predictions:

```{r coefficients-and-predictions}
head(coef(cv$scad$fit))
head(as.data.frame(prediction))
```

## Paths, tuning and interpretation

Use `gwr_ridge_path()`, `gwr_lasso_path()`, `gwr_en_path()`,
`gwr_scad_path()` or `gwr_mcp_path()` when the complete penalty path is needed.
The predictors passed to matrix interfaces exclude the intercept; the package
handles it separately. Predictor standardization is enabled by default and
coefficients are transformed back to the original predictor units.

```{r regularization-path}
path <- gwr_scad_path(x, train$y, coords = coords, k = 18,
  lambda = c(0.2, 0.08, 0.02), control = control)
path$lambda
```

`cv_gwr_penalized()` accepts an explicit `penalty` and returns CV diagnostics,
the selected tuning row and, with `refit = TRUE`, a fitted model. Dedicated
`cv_gwr_scad()` and `cv_gwr_mcp()` interfaces are also available. Training-fold
standardization and training-only neighbors prevent preprocessing and
neighborhood leakage. Hold folds fixed for a paired method comparison.

The fixed penalty grid in this example is specified before CV. An adaptive
bandwidth or mixing-parameter search requires an explicitly designed outer
search; this guide does not perform such a search. Examine convergence,
candidate eligibility and numerical diagnostics before interpreting results.
The default SCAD/MCP solver remains `coordinate`.

## Paper-specific source versions

This guide describes the current package. Reproducing the paper requires its
frozen simulation (0.4.0) and county-analysis (0.4.0.9006) source versions,
settings, inputs and software records. Use the accompanying reproduction
archive for those studies. This demonstration is intentionally small and is
not a substitute for either numerical experiment.
